%!TEX root = ../main.tex


\section{Chain Contractions and Discrete Morse Theory}\label{sec:contractions}

In this section we'll build the theoretical tools for computing Conley-filtered complexes.   We detail a computational version of the theory presented here in a later section.  We'll first review the tools for the chain complexes and the category $Ch(k)$ before proceeding to graded and filtered versions within the categories ${\bf GCC}$ and ${\bf LFC}$, respectively.  Much of the material may feel redundant, as we will port results from chain complexes to graded and filtered versions.


  In computational homological algebra, one often finds a simpler representative to compute homology from.  A model for this is the notion of {\em chain contraction}, which is a particular type of chain homotopy equivalence.  The notion also goes under the moniker {\em strong deformation retract}.  It can be found in Eilenberg and MacLane, homological perturbation theory and forms the basis for effective homology theory and algebraic Morse theory~\cite{}.

\begin{defn}
{\em A {\em chain contraction} is a diagram of chain complexes
\[
\xymatrixrowsep{0.03in}
\xymatrixcolsep{0.3in}
\xymatrix{
M  \ar[r]<3pt>^{\phi} & \ar[l]<3pt>^{\psi} C \ar@(u,r)^{\gamma}
}
\]
where $\phi,\psi$ are chain maps and $\gamma$ is a degree+1 map satisfying 
\[
\phi\psi = id_M\quad\text{ and }\quad \phi\psi = id_C-(\gamma d+d\gamma)
\]
}
\end{defn}

One often assumes the following identities, called the {\em side conditions}
\[
\gamma^2 = \gamma\phi = \psi\gamma = 0
\]
 
The side conditions may be assumed without loss of generality by replacing $\gamma$ with a particular function involving $\gamma$ and $d$, see~\cite{gug}. From the definition it is clear that $\phi$ is a monomorphism and $\psi$ is an epimorphism.  In applications, one calls $M$ the {\em reduced complex}.  When chain contractions arise from discrete-algebraic Morse theory it is called the {\em Morse complex}.  The point is that the reduced complex often has much smaller cardinality than $C$, leading to efficient computation of homology.  Contractions may be obtained from special degree +1 maps called splitting homotopies.

\begin{defn}
{\em
Let $(C,d)$ be a chain complex.  A {\em splitting homotopy} is a degree +1 map $h:C\to C$ such that $h^2=0$ and $h dh = h$.
}
\end{defn}

The conditions $d^2=h^2=0$ and $hdh = h$ ensure that $hd+dh$ is a projection.  Therefore $\pi=id_C-(hd+dh)$ is a projection onto the complementary subspace.  The image $(M,d^M)=(im(\pi),d|_{im(\pi)})$ is a subcomplex of $C$.  We have the following contraction:
\[
\xymatrixrowsep{0.03in}
\xymatrixcolsep{0.3in}
\xymatrix{
M  \ar[r]<3pt>^{i} & \ar[l]<3pt>^{\pi} C \ar@(u,r)^{h}
}
\]

We can calculate the differential $d^M$ via $$d\pi = d(id-(hd+dh)) = d-dhd + ddh = d-dhd$$

Finally, it is straightforward that the remaining side conditions $\gamma i=\pi\gamma = 0$ are easily verified. 

\begin{prop}\label{prop:cont:homiso}
Contractions and splitting homotopies are in bijective correspondence, up to isomorphism.
\end{prop}
\begin{proof}
Consider two contractions:
\[
\xymatrixrowsep{0.03in}
\xymatrixcolsep{0.3in}
\xymatrix{
M  \ar[r]<3pt>^{i} & \ar[l]<3pt>^{\pi} C \ar@(ul,ur)^{h} \ar[r]<3pt>^{\pi'}  & M'  \ar[l]<3pt>^{i'}
}
\] 

It is straightforward from the side conditions that the compositions $\pi\circ i'$ and $\pi'\circ i$ are inverses.  Therefore $M$ and $M'$ are chain isomorphic.
\end{proof}

\begin{ex}
Let $\cX$ be a cell complex and $(\cA,w:\cQ\to \cK)$ an acyclic partial matching.  By Proposition~\ref{prop:matchinghomotopy} there exists a unique splitting homotopy $\gamma$.  By Theorem~\ref{} we have a chain contraction:
\[
\xymatrixrowsep{0.03in}
\xymatrixcolsep{0.3in}
\xymatrix{
(C_\bullet(\cA),\partial^\cA)  \ar[r]<3pt>^{\phi} & \ar[l]<3pt>^{\psi} (C_\bullet(\cX),\partial^\cX) \ar@(ul,ur)^{\gamma}
}
\]
Notice that for $(M,d^M)$ given by the formulas above, we have
\[
(C_\bullet(\cA),\partial^\cA) \cong (M,d^M)
\]
via Proposition~\ref{prop:cont:homiso}.
\end{ex}


  We say a contraction is {\em strict} if the $M$ is cyclic.  We say a splitting homotopy $h$ is {\em perfect} if $d=dhd$. 

\begin{prop}
Strict contractions and perfect splitting homotopies are in bijective correspondence.
\end{prop}
\begin{proof}
If the contraction is strict then $di\pi = id^M\pi = 0$.  We have $i\pi = id_C-dh-hd$ and apply $d$ to each side to get $$0=d(i\pi) = d(id_C-dh-hd) = d-dhd$$

Conversely, if $\gamma$ is perfect then with $M=im(\pi)$ the differential $d^M$ is calculated as $$d^M=d-dhd=0$$ 

Therefore $M$ is cyclic and the contractions is strict.
\end{proof}

A perfect splitting homotopy thus implies $im(\pi)\cong H_\bullet (C)$.  This allows the homology to be read from the contraction without computation.  In addition, we have $d^Ci = id^M = 0$.  Therefore $i:M\to \ker d_C$ and $i(M)$ gives representatives for the homology in the original complex $C$. In the case of fields, perfect splittings always exist, see.  This implies a chain complex $C$ and its homology $H(C)$ always fit into a chain contraction. Moreover any chain contraction of homology, or more generally a cyclic complex, is minimal in the sense that the two complexes are isomorphic.

\begin{prop}\label{prop:cont:cyclic}
Let $(C,d)$ be a cyclic chain complex. For the contraction
\[
\xymatrixrowsep{0.03in}
\xymatrixcolsep{0.3in}
\xymatrix{
M  \ar[r]<3pt>^{i} & \ar[l]<3pt>^{\pi} C \ar@(u,r)^{h}
}
\]
We have $M\cong C$.
\end{prop}
\begin{proof}
We have $\pi \circ i  = id_M$.  If $C$ is cyclic then $d=0$ and $i\circ\pi = id_C-(hd+dh) = id_C$.
\end{proof}

In this sense, the homology $H(C)$ is the algebraic core of a chain complex and the minimal model for $C$ with respect to chain contractions.  This result will have analogues in the graded and filtered cases.

Finally, we should that chain contractions compose.  

\begin{prop}
Consider the sequence of contractions:
\[
\xymatrixrowsep{0.03in}
\xymatrixcolsep{0.35in}
\xymatrix{
M' \ar[r]<3pt>^{i'} & \ar[l]<3pt>^{\pi '} M \ar@(ul,ur)^{\gamma'}  \ar[r]<3pt>^{i} & \ar[l]<3pt>^{\pi} C \ar@(u,r)^{\gamma}
}
\]
Then there is a chain contraction 
\[
\xymatrixrowsep{0.03in}
\xymatrixcolsep{0.35in}
\xymatrix{
M''  \ar[r]<3pt>^{i''} & \ar[l]<3pt>^{\pi''} C \ar@(u,r)^{\gamma''}
}
\]
with the maps given by the formulas
\[
i'' = i \circ i ' \quad\quad \pi'' = \pi'\circ  \pi \quad\quad \Gamma = \gamma + i \circ \gamma '\circ \pi
\]

\end{prop}
\begin{proof}
Elementary computations show that 
\[
\pi''\circ i'' = id_{M''}\quad\text{and}\quad i'' \circ \pi'' = id_C-(d\Gamma + \Gamma d)
\]

The side conditions follow from the side conditions for $\gamma$ and $\gamma'$
\begin{enumerate}
\item  $\Gamma^2 = (\gamma+i\gamma'\pi)(\gamma+i\gamma'\pi) = \gamma^2 +  (\gamma i)\gamma'\pi + i\gamma'(\pi\gamma)  + i\gamma' (\pi i)\gamma'\pi = i (\gamma'\gamma' )\pi = 0$
\item $ \Gamma \circ i'' = (\gamma + i\gamma'\pi)(i\circ i') = (\gamma i)i'  + \pi' (\pi i)\gamma '\pi =  (\pi'\gamma) '\pi = 0$
\item $\pi''\circ \Gamma = (\pi'\pi)(\gamma+i\gamma'\pi) = \pi' (\pi\gamma) + \pi'( \pi i) \gamma'\pi = (\pi'\gamma' )\pi = 0$
\end{enumerate}

\end{proof}

An inductive argument gives the following result:


\begin{prop}\label{prop:cont:tower}
For a tower of chain contractions:
\[
\xymatrixrowsep{0.03in}
\xymatrixcolsep{0.35in}
\xymatrix{
M_n \ar[r]<3pt>^{i_n} & \ar[l]<3pt>^{\pi_n } M_{n-1} \ar@(ul,ur)^{\gamma_n} \ar[r]<3pt>^{i_{n-1}} & \ar[l]<3pt>^{\pi_{n-1}}  \ldots \ar[r]<3pt>^{i_1}  & \ar[l]<3pt>^{\pi_1 }M_0 \ar@(ul,ur)^{\gamma_1}  \ar[r]<3pt>^{i_0} & \ar[l]<3pt>^{\pi_0} C \ar@(u,r)^{\gamma_0}
}
\]

\begin{enumerate}
\item there is a chain contraction 
\begin{align}\label{dia:cont:tower}
\xymatrixrowsep{0.03in}
\xymatrixcolsep{0.35in}
\xymatrix{
M_n \ar[r]<3pt>^{i_{n+1}} & \ar[l]<3pt>^{\pi_{n+1}} C \ar@(u,r)^{\gamma_{n+1} }
}
\end{align}
with the maps given by the formulas
\[
i_{n+1} = \prod_{k=0}^n i_k \quad\quad \pi_{n+1} = \prod_{k=0}^n \pi_k \quad\quad \gamma = \sum_{k=0}^n i_k \gamma^k \pi_k
\]
\item $\gamma$ is a splitting homotopy and $\gamma$ is perfect if any $\gamma_i$ is perfect.

\end{enumerate}
\end{prop}
\begin{proof}
(1) follows from induction.  That $\gamma$ is a splitting homotopy follows since Diagram~\ref{dia:cont:tower} is a contraction.  If $\gamma_i$ is perfect, then $M_i$ is cyclic.  Thus $M_j$ is cyclic for any $j\geq i$ by Proposition~\ref{prop:cont:cyclic}. In particular $M_n$ is cyclic and Diagram~\ref{dia:cont:tower} is strict.  Therefore $\gamma$ is a perfect splitting homotopy.
\end{proof}






\subsection{Graded Contractions}

Filtered and graded versions of the theory are obtained by regarding the diagram in the appropriate category.  A {\em $\sP$-graded chain contraction} is a diagram in the category ${\bf GCC(\sP)}$
 \[
\xymatrixrowsep{0.03in}
\xymatrixcolsep{0.3in}
\xymatrix{
M^\oplus(\sP)  \ar[r]<3pt>^{\Phi} & \ar[l]<3pt>^{\Psi} C^\oplus(\sP)\ar@(ul,ur)^{\Gamma}
}
\]

 An $\sP$-graded chain contraction is {\em strict} if $M^\oplus(\sP)$ is a Conley complex.  A $\sP$-graded splitting homotopy is a degree+1 map $\Gamma:C^\oplus(\sP)\to C^\oplus(\sP)$ that is both a splitting homotopy and $\sP$-graded.    Again, one may define $\Pi=id_C-(\Gamma\Delta+\Delta\Gamma)$ and $M=im(\Pi)$.  Then $M$ is a subcomplex of $C$, $\Pi \circ i = id_M$ and $i\circ \Pi = id_C-(\Gamma\Delta+\Delta\Gamma)$.  
  
\begin{prop}\label{prop:grad:contract}
Let $C^\oplus(\sP)$ be a $\sP$-graded complex.  Let $\Gamma$ be a graded splitting homotopy and $\Pi = id_C-(\Gamma\Delta+\Delta\Gamma)$.  Define $M = im(\Pi)$.  Then $M$ is $\sP$-graded.
\end{prop}
\begin{proof}
Let $M_p = C_p\cap M$.  Then $$M=\bigoplus_{p\in \sP} M_p$$

Moreover $\Delta|_M$ is $\sP$-graded since $\Delta$ is $\sP$-graded.
\end{proof}

\begin{cor}
Graded splitting homotopies and $\sP$-graded chain contractions are in bijective corespondence.
\end{cor}


For $C^\oplus(\sP)$ a graded splitting homotopy $\Gamma\colon C^\oplus(\sP)\to C^\oplus(\sP)$ is {\em perfect} if for each $p$ the splitting homotopy $\Gamma_{pp}:C_p\to C_p$ is perfect, i.e. we have that $$\Delta_{pp} = \Delta_{pp}\Gamma_{pp}\Delta_{pp}$$  Such graded homotopies give rise to Conley complexes.

\begin{cor}
Perfect graded splitting homotopy and strict $\sP$-graded chain contractions are in bijective corespondence.
\end{cor}
\begin{proof}
A contraction implies the equation $$i\Delta^M\Pi = \Delta (i\circ \Pi) = \Delta(id_C-\Gamma\Delta-\Delta\Gamma) = \Delta-\Delta\Gamma\Delta$$  Since these maps are $\sP$-graded we have $$0=i_{pp}\Delta^M_{pp}\Pi_{pp}= (i\Delta^M\Pi)_{pp}  =(\Delta-\Delta\Gamma\Delta)_{pp}= \Delta_{pp}-\Delta_{pp}\Gamma_{pp}\Delta_{pp}$$

Conversely, let $\gamma$ be a perfect graded splitting homotopy.  The differential on $M=im(\Pi)$ is calculated as $\Delta^M = \Delta-\Delta\Gamma\Delta$.  Since the maps $\Delta$ and $\Gamma$ are $\sP$-graded (upper triangular), we have $$\Delta^M_{pp} = (\Delta-\Delta\Gamma\Delta)_{pp} = \Delta_{pp}-\Delta_{pp}\Gamma_{pp}\Delta_{pp} = 0$$

\end{proof}


We can now show that Conley complexes are minimal with respect to chain contractions.  Again, this is an elementary observation, but it mirrors the above Proposition~\ref{prop:cont:cyclic}.

\begin{prop}
Let $C^\oplus(C)$ be a Conley complex.  Any contraction
 \[
\xymatrixrowsep{0.03in}
\xymatrixcolsep{0.3in}
\xymatrix{
M^\oplus(\sP)  \ar[r]<3pt>^{\Phi} & \ar[l]<3pt>^{\Psi} C^\oplus(\sP) \ar@(u,r)^{\Gamma}
}
\]
is strict. Moreover $M^\oplus(\sP)\cong C^\oplus(\sP)$.
\end{prop}
\begin{proof}
The differential on $M$ is calculated by $\Delta^M = \Delta-\Delta\Gamma\Delta$.  We have 
\[
\Delta^M_{pp} = (\Delta-\Delta\Gamma\Delta)_{pp} = \Delta_{pp}-\Delta_{pp}\Gamma_{pp}\Delta_{pp} = 0
\]
There $M^\oplus$ is a Conley complex.  Since $i$ and $\Pi$ is a chain homotopy equivalence, invoking Proposition~\ref{prop:grad:cmiso} shows that $M^\oplus(\sP)$ and $C^\oplus(\sP)$ are graded chain isomorphic. 
\end{proof}

For a tower of graded chain contractions, we have the following.  This is analogous to Proposition~\ref{prop:cont:tower}.

\begin{prop}
For a tower of $\sP$-graded chain contractions:
\[
\xymatrixrowsep{0.03in}
\xymatrixcolsep{0.35in}
\xymatrix{
M_n^\oplus(\sP) \ar[r]<3pt>^{i_n} & \ar[l]<3pt>^{\pi_n } M_{n-1}^\oplus(\sP) \ar@(ul,ur)^{\gamma_n} \ar[r]<3pt>^{i_{n-1}} & \ar[l]<3pt>^{\pi_{n-1}}  \ldots \ar[r]<3pt>^{i_1}  & \ar[l]<3pt>^{\pi_1 }M_0 \ar@(ul,ur)^{\gamma_1}  \ar[r]<3pt>^{i_0} & \ar[l]<3pt>^{\pi_0} C^\oplus(\sP) \ar@(u,r)^{\gamma_0}
}
\]

\begin{enumerate}
\item there is a chain contraction 
\begin{align}\label{dia:cont:tower}
\xymatrixrowsep{0.03in}
\xymatrixcolsep{0.35in}
\xymatrix{
M_n^\oplus(\sP) \ar[r]<3pt>^{i_{n+1}} & \ar[l]<3pt>^{\pi_{n+1}} C^\oplus(\sP) \ar@(u,r)^{\gamma_{n+1} }
}
\end{align}
with the maps given by the formulas
\[
i_{n+1} = \prod_{k=0}^n i_k \quad\quad \pi_{n+1} = \prod_{k=0}^n \pi_k \quad\quad \gamma = \sum_{k=0}^n i_k \gamma^k \pi_k
\]
\item $\gamma$ is a splitting homotopy and $\gamma$ is perfect if any $\gamma_i$ is perfect.

\end{enumerate}
\end{prop}



\subsection{Filtered Contractions}
 
An {\em $\sL$-filtered chain contraction} is a diagram in the category $\bCF$ 

\[
\xymatrixrowsep{0.03in}
\xymatrixcolsep{0.3in}
\xymatrix{
m  \ar[r]<3pt>^{\phi} & \ar[l]<3pt>^{\psi} f \ar@(u,r)^{h}
}
\]

where $\psi\phi = id_M$ and $\phi\psi = id_C-(hd+dh)$.  Here  $M$ and $C$ are $\sL$-filtered chain complexes, $f$ and $g$ are $\sL$-filtered chain maps and $h$ is an $\sL$-filtered degree+1 map.  An $\sL$-filtered chain contraction is {\em strict} if $(M,d)$ is a Conley filtering.  The existence proof of~\cite{salamon}  furnish a filtered chain contraction.  

A filtered splitting homotopy is a degree+1 map $h:C\to C$ that is both a splitting homotopy and filtered with respect to $\sL$.  Again, one may define $\pi=id_C-(hd+dh)$ and $M=im(\pi)$.  Then $M$ is a subcomplex of $C$, $\pi i = id_M$ and $i\pi = id_C-(hd+dh)$.  The next result shows that $M$ is $\sL$-filtered. 

\begin{prop}
Let $f:\sL\to Sub(C)$ be an $\sL$-filtered complex.  Let $h$ be a filtered splitting homotopy and $\pi = id-(hd+dh)$.  Define $M=im(\pi)$ and $m:\sL\to Sub(M,d)$ as the map $$L\ni q\rightsquigarrow  \pi(f(q))\in Sub(M)$$

Then $m:L\to Sub(M,d)$ is an $\sL$-filtered complex.
\end{prop}
\begin{proof}
We begin with showing that $m(p \wedge q) = m(p)\wedge m(q)$.  We have that $$m(p\wedge q) = \pi(f(p\wedge q)) = \pi(f(p)\wedge f(q))$$

We must show that $\pi(f(p)\cap f(q)) = \pi(f(p))\cap \pi(f(q))$.  It is elementary set theory that $\pi(f(p)\cap f(q))\subseteq \pi(f(p))\cap \pi(f(q))$.  Now let $x\in \pi(f(p))\cap \pi(f(q))$.  Since $\pi(i(x))=x$ it suffices to show that $i(x)\in f(p)\cap f(q)$.  By definition there are $y\in f(p)$ and $y'\in f(q)$ such that $\pi(y) = x = \pi(y')$.  We have that $$i(x) = i(\pi(y)) = (id_C-hd - dh)(y)$$

The map on the right hand side is filtered since $h$ and $d$ are filtered.  Thus $i(x)\in f(p)$.  Similarly, $i(x)\in f(q)$.  Therefore $i(x)\in h(p)\cap h(q)$.  Finally, it is a straightforward consequence of the linearity of $\pi$ that $\pi(f(p\vee q)) = \pi(f(p))+\pi(f(q))$.


\end{proof}

\begin{cor}\label{cor:filt}
Filtered splitting homotopies and $\sL$-filtered chain contractions are in bijective correspondence.
\end{cor}

For $f\in \bCF(L,k)$ a filtered splitting homotopy $h:f\to f$ is {\em perfect} if for each $q\in J(L)$ the induced map $$h:C_q/C_{Pred(q)}\to C_q/C_{Pred(q)}$$ is perfect.  Such filtered homotopies give rise to Conley filterings.

\begin{cor}
Perfect filtered splitting homotopies are in bijective correspondence with strict $\sL$-filtered chain contractions.
\end{cor}
\begin{proof}
 Consider $m:\sL\to Sub(M,d^M)$ of the $\sL$-filtered contraction guaranteed by Corollary~\ref{cor:filt}.  If $h$ is perfect the differential $d^M$ must must obey $d^M(M_q) = (d-dhd)M_q\subseteq M_{Pred(q)}$.  Therefore $m:\sL\to Sub(M,d^M)$ is a Conley filtered.
\end{proof}

The purpose of our computational section is to show how to furnish perfect filtered splitting homotopies.  


\begin{prop}
Let $f\colon\sL\to Sub(C,d)$ be an $\sL$-Conley filtered complex.  Any filtered chain contraction
\[
\xymatrixrowsep{0.03in}
\xymatrixcolsep{0.3in}
\xymatrix{
m  \ar[r]<3pt>^{\phi} & \ar[l]<3pt>^{\psi} f \ar@(u,r)^{h}
}
\]
is strict.  Moreover $m$ and $f$ are filtered chain isomorphic.
\end{prop}
\begin{proof}
 The boundary operator is calculated as $d^M= d-dhd$.  Since $f$ is Conley-filtered  we have $d^M(M_q) = (d-dhd)M_q\subseteq M_{Pred(q)}$.  Thus $m$ is Conley-filtered.  Finally, $m$ and $f$ are filtered chain isomorphic by Proposition~\ref{prop:filt:cmiso}.
\end{proof}

For a tower of $\sL$-filtered chain contractions, we have:

\begin{prop}
For a tower of $\sL$-filtered chain contractions:
\[
\xymatrixrowsep{0.03in}
\xymatrixcolsep{0.35in}
\xymatrix{
m_n \ar[r]<3pt>^{i_n} & \ar[l]<3pt>^{\pi_n } m_{n-1} \ar@(ul,ur)^{\gamma_n} \ar[r]<3pt>^{i_{n-1}} & \ar[l]<3pt>^{\pi_{n-1}}  \ldots \ar[r]<3pt>^{i_1}  & \ar[l]<3pt>^{\pi_1 } m_0 \ar@(ul,ur)^{\gamma_1}  \ar[r]<3pt>^{i_0} & \ar[l]<3pt>^{\pi_0} f \ar@(u,r)^{\gamma_0}
}
\]

\begin{enumerate}
\item there is a chain contraction 
\begin{align}\label{dia:cont:tower}
\xymatrixrowsep{0.03in}
\xymatrixcolsep{0.35in}
\xymatrix{
m_n \ar[r]<3pt>^{i_{n+1}} & \ar[l]<3pt>^{\pi_{n+1}} f \ar@(u,r)^{\gamma_{n+1} }
}
\end{align}
with the maps given by the formulas
\[
i_{n+1} = \prod_{k=0}^n i_k \quad\quad \pi_{n+1} = \prod_{k=0}^n \pi_k \quad\quad \gamma = \sum_{k=0}^n i_k \gamma^k \pi_k
\]
\item $\gamma$ is a splitting homotopy and $\gamma$ is perfect if any $\gamma_i$ is perfect.

\end{enumerate}
\end{prop}


\subsection{Graded Morse Theory}

A graded version of discrete Morse theory may be done straightforwardly.   Let $f:X\to \sP$ be a $\sP$-graded cell complex.  We say that $(A,w)$ is a {\em graded acyclic partial matching} if it satisfies $w(Q)=K$ only if $K,Q\in X_p$ for some $p\in P$.  That is, matchings may only occur in the same fiber of the valuation.  This idea can be found many places in the literature, for instance, see~\cite{mn} and~\cite[Patchwork Theorem]{koz}.

\begin{prop}
Let $f\colon \cX \to \sJ(\sL)$ be a graded cell complex and $(\cA,w)$ a graded acyclic matching.  Then the associated splitting homotopy $\Gamma$ for $(C^\oplus(\sJ(\sL)),\Delta)$ is $\sP$-graded and fits into a graded chain contraction.
\end{prop}
\begin{proof}
By Proposition~\ref{prop:matchinghomotopy} there is a splitting homotopy $\Gamma:C(X)\to C(X)$ associated to the matching $(\cA,w)$.  Let $p\in J(L)$.  Consider $(\cA_p,w_p)$ the matching restricted to the fiber $X_p = f^{-1}(p)$.  We have $$\cA_p = \cA\cap \cX_p\quad\quad \quad \quad  w_p:\cQ\cap X_p\to \cK\cap \cX_p$$

By assumption this is an acyclic partial matching on the fiber $X_p$.  By Proposition~\ref{prop:matchinghomotopy} there is a unique splitting homotopy $\gamma_p:C(X_p)\to C(X_p)$.    Therefore we have $$C(X) = \bigoplus_{p\in \sP} C(X_p) \quad\quad \Gamma_{pp} = \gamma_p$$ 

%Thus $\Gamma$ is diagonal, and in particular a graded-splitting homotopy.
{\bf still needs proof that $\Gamma$ is graded, }

Since $\Gamma$ is graded by Proposition~\ref{prop:grad:contract} there is an associated contraction
\[
\xymatrixrowsep{0.03in}
\xymatrixcolsep{0.3in}
\xymatrix{
A^\oplus(\sP)  \ar[r]<3pt>^{\Phi} & \ar[l]<3pt>^{\Psi} C^\oplus(\sP) \ar@(ul,ur)^{\Gamma}
}
\]

\end{proof}



\subsection{Connection Matrix Algorithm}

In this section we introduce the algorithm for computing a connection matrix based on the Morse theory described above.

{\bf Algorithm}
\begin{enumerate}
\item Given a graded cell complex $f_0\colon \cX\to \sJ(\sL)$ as input
\item {\bf do}
\item Apply~\cite[Algorithm 3.6]{focm} to the fibers $\{X_q\colon X_q = f^{-1}(q)\}$ to produce a graded acyclic partial matching $(\cA,w:\cQ\to \cK)$
\item Apply~\cite[Algorithm 3.12]{focm} to produce a graded splitting homotopy $\Gamma:C^\oplus(\sP)\to C^\oplus(\sP)$ and graded chain contraction
\[
\xymatrixrowsep{0.03in}
\xymatrixcolsep{0.3in}
\xymatrix{
A^\oplus(\sP)  \ar[r]<3pt>^{\Phi} & \ar[l]<3pt>^{\Psi} C^\oplus(\sP) \ar@(ul,ur)^{\Gamma}
}
\]
\item {\bf while $|\cA|<|\cX|$}
\end{enumerate}


\begin{thm}
The above algorithm terminates, i.e. after finitely many iterations the sequence $(f_n)$ stabilizes to a final $L$-filtered complex $f_\infty$.  Moreover, $f_\infty$ is a Conley complex.
\end{thm}
\begin{proof}

\end{proof}





